### 2024-08-23 paper/CPP/paper.tex
Finish section on Explain_Simple

\documentclass[
  sigplan,
  10pt,
  anonymous,
  review,
  ]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\usepackage[british]{babel}

\usepackage[capitalise]{cleveref}

\usepackage{mathpartir}

% Record brackets
\usepackage{stmaryrd}

\usepackage{xcolor}

\usepackage{tikz}
\usepackage{pgfplots}

\usepackage{enumitem}

\usepackage{array}
\usepackage{multirow}

\usepackage{listings}
\definecolor{isarblue}{HTML}{006699}
\definecolor{isargreen}{HTML}{009966}
\lstdefinelanguage{isabelle}{%
    keywords=[1]{type_synonym,typedef,datatype,fun,function,abbreviation,definition,proof,lemma,theorem,corollary,inductive},
    keywordstyle=[1]\bfseries\color{isarblue},
    keywords=[2]{where,assumes,shows,and,fixes,defines},
    keywordstyle=[2]\bfseries\color{isargreen},
    keywords=[3]{if,then,else,case,of,SOME,let,in,O},
    keywordstyle=[3]\color{isarblue},
}
\lstset{%
  language=isabelle,
  escapeinside={&}{&},
  columns=\lst@ifdisplaystyle{fullflexible}\else{fixed}\fi,,
  extendedchars,
  basewidth={0.5em,0.45em},
  basicstyle=\ttfamily,
  mathescape,
}
\makeatletter
\lst@AddToHook{OnEmptyLine}{\vspace{-0.4\baselineskip}}
\makeatother
\newcommand{\lstinlinew}[2][]{\lstinline[#1]{#2}}
% This is a hack to allow escaping inside lstinline (https://tex.stackexchange.com/questions/43526/escaping-in-lstinline)
\usepackage{etoolbox}
\makeatletter
\patchcmd{\lsthk@TextStyle}{\let\lst@DefEsc\@empty}{}{}{\errmessage{failed to patch}}
\makeatother

\usepackage{calc}

\newcommand{\textover}[3][l]{%
 % #1 is the alignment, default l
 % #2 is the text to be printed
 % #3 is the text for setting the width
 \makebox[\widthof{#3}][#1]{#2}%
}

\hyphenation{Isa-belle}

% For arXiv: LuaLatex on Arxiv: https://tex.stackexchange.com/questions/372154/lualatex-how-to-produce-pdf-acceptable-by-arxiv 
%\hypersetup{%
%  pdfcreator = {},
%  pdfproducer = {}
%}
%\pdfvariable suppressoptionalinfo \numexpr 1+2+4+8+16+32+64+128+256+512 \relax

% For arXiv: Remove copyright information
% \setcopyright{none}
% \renewcommand\footnotetextcopyrightpermission[1]{}

\usepackage{todonotes}

% Autoref names
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}

% Notation
\newcommand{\TransP}{\bigtriangledown}

\begin{document}

% For arXiv: Remove conference/author in the header
% \pagestyle{plain}

\setcounter{tocdepth}{1}

\title{Simplified and Verified: A Fresh Look at a Proof-Producing Union-Find Algorithm}
\author{Lukas Stevens}
\orcid{0000-0003-0222-6858}
\affiliation{%
  \institution{Technical University of Munich}
  \department{Department of Informatics}
  \streetaddress{Boltzmannstr. 3}
  \city{Garching}
  \postcode{85748}
  \country{Germany}
}
\email{lukas.stevens@in.tum.de}

\begin{abstract}
  TODO
\end{abstract}

\maketitle

\listoftodos{}

\section{Introduction}
\begin{itemize}
  \item Based on \cite{congcl_proofs}
\end{itemize}
\subsection{Contributions}

\subsection{Related Work}
\begin{itemize}
  \item Congruence closure in Coq by Corbineau
\end{itemize}

\subsection{Notation}
Isabelle/HOL~\cite{isabelle} conforms to everyday mathematical notation for the most part.
We establish notation and in particular some essential data types together with their primitive operations that are specific to Isabelle/HOL.

We write \lstinline!t :: 'a! to specify that the term \lstinline!t! has the type \lstinline!'a! and \lstinline!'a $\Rightarrow$ 'b! for the space of total functions from type \lstinline!'a! to type \lstinline!'b!.

Sets with elements of type \lstinline!'a! have the type \lstinline!'a set!.
The cardinality of a set \lstinline!A! is denoted by \lstinline!card A! and the image of \lstinline!A! under \lstinline!f! by \lstinline!f ` A!.

We use \lstinline!'a list! to describe the type of lists, which are constructed using the empty list \lstinline![]! constructor or the infix cons constructor \lstinline!#!, and are appended with the infix operator \lstinline!@!.
The function \lstinline!set! converts a list into a set.
\todo[inline]{Indexing operator}
\todo[inline]{Relation type synonym, Field of a relation}

We remark that $\longleftrightarrow$ is equivalent to \lstinline!$=$! on the type of Booleans \lstinline!bool! and \lstinline!$\equiv$! is definitional equality of the meta-logic of Isabelle/HOL, which is called Isabelle/Pure.
Meta-implication is denoted by \lstinline!$\Longrightarrow$! and a chain of implications
\begin{lstlisting}
  A$_\text{1}$ $\Longrightarrow$ $\cdots$ $\Longrightarrow$ A$_\text{k}$ $\Longrightarrow$ C
\end{lstlisting}
can be abbreviated by 
\begin{lstlisting}
  $\llbracket$ A$_\text{1}$;$\,\ldots\,$;A$_\text{k}$ $\rrbracket$ $\Longrightarrow$ C.
\end{lstlisting}

\section{Basic Union-Find}
\subsection{Background}
Given a set of $n$ elements $A = \{a_1, \ldots, a_n\}$, the union-find data structure keeps track of a partition of $A$ into disjoint sets $A_1, \ldots, A_k$, i.e.\ $A = A_1 \uplus \cdots \uplus A_k$.

We initialise the data structure by partitioning $A$ into singleton sets of elements,
so we have that $A = \{a_1\} \uplus \cdots \uplus \{a_n\}$.
Those sets are merged by subsequent \lstinline!union! operations where \lstinline!union $a_i$ $a_j$! merges the set containing $a_i$ with the one that contains $a_j$.
Each set in the partition contains one particular element that serves as its representative.
We will denote the representative of an element $a$ in the union-find data structure $l$ as \lstinline!rep_of $l$ $a$!.
Accordingly, two elements have the same representative exactly when they belong to the same set in the partition.
For any element $a_i$, the \lstinline!find! operation just returns its representative \lstinline!rep_of $l$ $a_i$!.

One application of this algorithm is to maintain the equivalence closure of equations, which is the partial equivalence relation formed by those equations.
To this end, we initialize the union-find data structure with $n$ elements for a given set of variables $v_1, \ldots, v_n$ and perform a \lstinline!union! between $v_i$ and $v_j$ for each equation $v_i = v_j$.
This results in a partition of the variables into equivalence classes.

According to \citeauthor{unionfind_tarjan}~\cite{unionfind_tarjan}, this data structure can be implemented as a forest of rooted trees where each tree represents an equivalence class.
The edges of the trees are directed towards their root, which is the representative of the corresponding equivalence class.
To keep this invariant, we initialise the forest with $n$ nodes but without any edges and for every \lstinline!union! of $a_i$ and $a_j$ we add a directed edge from \lstinline!rep_of $l$ $a_i$! to \lstinline!rep_of $l$ $a_j$! to the forest.

We encode such a forest as a list $l$ of length $n$, where at each index $i$ of the list, we save the parent of the element $i$.
Therefore, we denote the parent of an element $a$ with \lstinline|$l$ ! $a$|.
If $a$ is a root and thus does not have a parent, then the list contains the element $a$ itself at the index $a$, i.e., \lstinline|$l$ ! $a$ = $a$|.

\todo[inline]{Mention optimisations?}

\subsection{In Isabelle/HOL}
The union-find algorithm is formalised in Isabelle/HOL by \citeauthor{unionfind_isabelle}~\cite{unionfind_isabelle}.
The code can be found in an entry~\cite{union_find_afp} of the \emph{Archive of Formal Proofs} (AFP).\footnote{The code is in the theory file \texttt{Examples/Union\_Find.thy}.}
Later, \citeauthor{union_find_imp_hol}~\cite{union_find_imp_hol} also formalised the running time.\todo{Check citations}

The formalisation defines a function \lstinline|rep_of|, which, as described in the previous section, follows the parent pointers until we arrive the root, where the parent pointer is self-referential.
\begin{lstlisting}[columns=fullflexible]
function rep_of :: nat list $\Rightarrow$ nat $\Rightarrow$ nat where
rep_of l a =
  if l ! a = a then a else rep_of l (l ! a)
\end{lstlisting}
Looking closely at this definition, we see that this function is only well-defined for some inputs \lstinline|l| and \lstinline|a|:
for every element \lstinline|a < length l|, its parent must be in the list, i.e.\ we must have \lstinline|l ! a < length l|, and the parent pointers must be cycle-free in order for the function to terminate.
Functions in Isabelle/HOL must be total, so the \lstinline|function| command introduces a constant
\begin{lstlisting}
rep_of_dom :: (nat list $\times$ nat) $\Rightarrow$ bool  
\end{lstlisting}
that characterises the inputs for which \lstinline|rep_of| terminates.
Then, it adds \lstinline|rep_of_dom (l, a)| as a premise to the defining equation of \lstinline|rep_of|. 
The intuition above is cast into a predicate \lstinline|ufa_invar| that defines such well-formed lists \lstinline|l|.
\begin{lstlisting}
ufa_invar l = $\forall$a < length l. rep_of_dom (l, a)
              $\land$ l ! a < length l
\end{lstlisting}

Building on this formalisation, we encapsulate this union-find data structure into an abstract data type (ADT) \lstinline|ufa| by defining
\begin{lstlisting}
typedef ufa = {uf. ufa_invar uf}
\end{lstlisting}
By lifting operations on the underlying list to the ADT using Isabelle's lifting infrastructure~\cite{lifting_transfer}, we obtain the operations
\begin{itemize}
  \item \lstinline!ufa_$\alpha$ :: ufa $\Rightarrow$ nat rel!,
  \item \lstinline!ufa_rep_of :: ufa $\Rightarrow$ nat $\Rightarrow$ nat!, and
  \item \lstinline!ufa_union :: ufa $\Rightarrow$ nat $\Rightarrow$ nat $\Rightarrow$ ufa!,
\end{itemize}
where \lstinline!nat rel = (nat $\times$ nat) set!.
The intended meaning of these operations is the following:
\begin{itemize}
  \item \lstinline!ufa_$\alpha$ uf! is the partial equivalence relation represented by \lstinline!uf!,
  \item \lstinline!ufa_rep_of uf x! is the representative of the equivalence class that \lstinline!x! belongs to, and
  \item \lstinline!ufa_union uf x y! returns a union-find data structure where the equivalence classes of \lstinline!x! and \lstinline!y! are merged.
    This is implemented by updating the underlying list at index \lstinline!rep_of l x! with \lstinline|rep_of l y|.
\end{itemize}
More rigorously, the above operations fulfil the following properties:\todo{Spacing between lemmas,defs}
\begin{lstlisting}
lemma ufa_rep_of uf x = ufa_rep_of uf y $\longleftrightarrow$
  (x, y) $\in$ ufa_$\alpha$ uf

lemma ufa_$\alpha$ (ufa_union uf x y) =
  per_union (ufa_$\alpha$ uf) x y
\end{lstlisting}
where \lstinline!per_union R x y!\todo{Definition with \lstinline|trancl|?} is defined as the equivalence relation that results from merging the respective equivalence classes in the relation \lstinline|R| that \lstinline!x! and \lstinline!y! belong to.

But what happens if \lstinline!x! or \lstinline!y! is not an element of the partial equivalence relation \lstinline|R|?\todo{Define element of per.}
In that case, the equivalence relation is unchanged, which means that \lstinline|per_union R x y = R|.
This, however, can be seen as a misuse of the union-find data structure, since we initialise it with a fixed set of elements \lstinline|A| and expect the user to only work with these elements.
Therefore, we introduce the following definitions that characterise valid union(s) with regard to the initial set \lstinline|A|.
\begin{lstlisting}
definition valid_union uf a b =
  a $\in$ Field (ufa_$\alpha$ uf) $\land$ b $\in$ Field (ufa_$\alpha$ uf)
definition valid_unions uf us =
  $\forall$(x, y) $\in$ set us. valid_union uf x y
\end{lstlisting}

\section{Simple Certifying Union-Find Algorithm}
\begin{figure}
  \begin{gather*}
    \inferrule*[right=Assm]{\hbox{\lstinline|i < length us|} \\ \hbox{\lstinline|us ! i = (x, y)|}}{\hbox{\lstinline|us $\ \vdash_=\ $ AssmP i : (x, y)|}} \\[0.5em]
    \inferrule*[right=Refl]{ }{\hbox{\lstinline|us $\ \vdash_=\ $ ReflP x : (x, x)|}} \\[0.5em]
    \inferrule*[right=Sym]{\hbox{\lstinline|us $\ \vdash_=\ $ p : (x, y)|}}{\hbox{\lstinline|us $\ \vdash_=\ $ SymP p : (y, x)|}} \\[0.5em]
    \inferrule*[right=Trans]{\hbox{\lstinline|us $\ \vdash_=\ $ p1 : (x, y)|} \\ \hbox{\lstinline|us $\ \vdash_=\ $ p2 (y, z)|}}{\hbox{\lstinline|us $\ \vdash_=\ $ p1 $\ \TransP\ $ p2 : (x, z)|}}
  \end{gather*}
  \caption{The data type \lstinline|eq_prf| of certificates and its corresponding system of inference rules $\vdash_=$.\label{fig:eq_prf}}
\end{figure}
Building on the union-find ADT from the previous section, we now develop a simple \lstinline|explain| operation that,
for a given list of equations \lstinline|us|, takes two elements \lstinline|x| and \lstinline|y| and produces a certificate that \lstinline|x = y| modulo \lstinline|us|.
The certificate is given in terms of a data type \lstinline|eq_prf| with its corresponding system $\vdash_=$ of inference rules as seen in \autoref{fig:eq_prf}.
As expected, we have inference rules that utilise the reflexivity, symmetry, and transitivity of equality as well as an assumption rule.
To improve readability, we use the infix operator $\bigtriangledown$ to denote the proof term for transitivity.

We remark that one can not prove disequality with this proof system.
In that sense, we are working with an open-world assumption.

We prove that $\vdash_=$ is sound and complete with respect to the equivalence relation induced by \lstinline|us|, i.e.\ the equivalence closure of \lstinline|us|.
In Isabelle, we define\todo{Introduce inverse and rtrancl?}
\begin{lstlisting}
definition symcl r = r $\cup$ r$^{-1}$
definition equivcl r = (symcl r)$^*$
\end{lstlisting}
and prove
\begin{lstlisting}
lemma $\vdash_=$_sound:
assumes as $\vdash_=$ p : (x, y)
shows (x, y) $\in$ equivcl (set as)

lemma $\vdash_=$_complete:
assumes (x, y) $\in$ equivcl (set as)
shows $\exists$p. as $\vdash_=$ p : (x, y)
\end{lstlisting}

\begin{figure*}
  \centering
  \begin{lstlisting}
function explain :: (nat $\times$ nat) list $\Rightarrow$ nat $\Rightarrow$ nat $\Rightarrow$ nat eq_prf where
  explain [] x _ = ReflP x
| explain (us @ [(a, b)]) x y =
    let uf = ufa_unions uf_init us; a_b_P = AssmP (length us)
    in if ufa_rep_of uf x = ufa_rep_of uf y then explain us x y
      else if ufa_rep_of uf x = ufa_rep_of uf a then explain us x a $\TransP$ a_b_P $\TransP$ explain us y b
      else explain us x b $\TransP$ SymP a_b_P $\TransP$ explain us a y

definition explain_partial us x y =
  if (x, y) $\in$ equivcl (set us) then Some (explain us x y) else None
  \end{lstlisting}
  \caption{A simple implementation of the \lstinline|explain| operation. It operates on an arbitrary initial union-find data structure \lstinline|uf_init|.\label{fig:explain}}
\end{figure*}
Although \lstinline|explain| takes a list of equations \lstinline|us|, its implementation actually works with an instance \lstinline|uf| of the union-find ADT that results from applying \lstinline|us| as unions to an initial union-find data structure \lstinline|uf_init|.
Accordingly, we define
\begin{lstlisting}
definition ufa_unions =
  foldl ($\lambda$uf (x, y). ufa_union uf x y)
\end{lstlisting}
so we have \lstinline|uf = ufa_unions uf_init us| in terms of the constants above.

At last, we are ready to implement the \lstinline|explain| operation as depicted in \autoref{fig:explain}.\todo{explain with Cons instead of snoc}
The algorithm works with the assumption that the given elements \lstinline|x| and \lstinline|y| are equal modulo \lstinline|us|.
Furthermore, we assume that each element of the initial data structure \lstinline|uf_init| is equal to itself
and that \lstinline|us| are valid unions with respect to \lstinline|uf_init|.
In formulae, we assume
\begin{itemize}
  \item \lstinline|ufa_$\alpha$ uf_init $\subseteq$ Id|, and
  \item \lstinline|valid_unions uf_init us|
\end{itemize}
in what follows.

If the list of unions is empty, then \lstinline|x| and \lstinline|y| must actually be equal which we certify with \lstinline|ReflP x|.

Otherwise, the list of unions must be equal to \lstinline|us @ [(a, b)]| for some \lstinline|us|, \lstinline|a|, and \lstinline|b|.
We distinguish between two cases:
\begin{itemize}
  \item The elements \lstinline|x| and \lstinline|y| are already equal modulo \lstinline|us|, so we proceed recursively with \lstinline|us|.
  \item In the case that the equation \lstinline|a = b| is necessary for \lstinline|x = y| to hold, we either have \lstinline|x = a| and \lstinline|b = y| or \lstinline|x = b| and \lstinline|a = y| modulo \lstinline|us|.
    Assuming that the former holds --- the other case is similar --- we recursively construct the certificates for \lstinline|x = a| and \lstinline|b = y|.
    Together with the fact that \lstinline|a = b|, which we prove by assumption, we obtain \lstinline|x = y| by transitivity.
\end{itemize}

We condense the above intuition into the completeness theorem below.
\begin{lstlisting}
theorem explain_complete:
assumes (x, y) $\in$ equivcl (set us)
shows us $\vdash_=$ explain uf_init us x y : (x, y)
\end{lstlisting}

The \lstinline|explain| function is not sound, though.
This is because it always returns a certificate, even if \lstinline|x| and \lstinline|y| are not equal modulo \lstinline|us|.
To account for this case, we wrap \lstinline|explain| into a partial function \lstinline|explain_partial| (cf.\ \autoref{fig:explain}) that fails if \lstinline|x = y| is not provable.
Soundness and completeness can then be lifted from the soundness of $\vdash_=$ and the completeness of \lstinline|explain|, respectively.
At first glance, it might seem vacuous to define \lstinline|explain_partial| using \lstinline|equivcl|;
however, \lstinline|equivcl| can actually be implemented using union-find operations as the following lemma demonstrated.
\begin{lstlisting}
lemma
defines uf = ufa_unions uf_init us
shows (x, y) $\in$ equivcl (set us) $\leftrightarrow$ x = y $\lor$
  x $\in$ Field (ufa_$\alpha$ uf) $\land$ y $\in$ Field (ufa_$\alpha$ uf) $\land$
  ufa_rep_of uf x = ufa_rep_of uf y
\end{lstlisting}
\todo[inline]{Improve explanation above? Segue?}

\begin{itemize}
  \item Inspired by formalisation by Lammich~\cite{unionfind_isabelle}, in AFP~\cite{union_find_afp}
  \item Refinements with ranks and int ranks
\end{itemize}

\subsection{Data Structure to enable explain}

\subsection{Union}

\subsection{Explain}

\subsection{Soundness and Completeness}

\section{Deriving an Efficient Implementation}
\begin{itemize}
  \item Refinements with ranks and ints
  \item Helper Functions
  \item Locales
  \item Induction principle and code equation
\end{itemize}

\section{Refinement to Imperative HOL}
\begin{itemize}
  \item Refining ufri
  \item Dynamic Array inspired by Auto2 Imperative HOL
  \item Refining explain operations
\end{itemize}

\section{Tactic}
Proof of concept tactic

\section{Conclusion and Future Work}

\begin{acks}
  TODO
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sources}

\end{document}