### 2025-07-25 paper/CADE/presentation/simple.tex

% TODO: remove pedantic step
% TODO: time travel joke? Use clock already?
\newcommand<>\gray[1]{%
  \textcolor#2{TUMLightGray}{#1}%
}
\begin{frame}{The simple algorithm: example}
\textbf{Input:}
\foreach \al/\g [count=\i] in {{18-19,34-35}/{36-},{16-17,28-33}/{18-20,34-},{14-15,25-27}/{16-20,28-},{12-13,23-24}/{14-20,25-},{10-11,21-22}/{12-20,23-},{1-9}/{10-}} {
  \action<uncover@1-|alert@\al|gray@\g>{$\underbrace{\eqN{\i}}_\i$}
}
\vfill
\begin{columns}
\begin{column}{0.3\textwidth}
\begin{forest}
  for tree={
    math content,
    draw,
    circle,
    l=1.6cm,
    s=1.4cm,
    edge={<-}
  }
  [\varN{2}
    [\varN{1}, edge={alert on={19-20,35}, gray on={36-}}, edge label={node[pos=0.6, above left] {$1$}}
      [\varN{7}, edge={alert on={11,22}, gray on={12-20,23-}}, edge label={node[midway, left] {$5$}}]]
    [\varN{6}, edge={alert on={13,24}, gray on={14-20,25-}}, edge label={node[pos=0.6, right] {$4$}}]
    [\varN{5}, edge={alert on={3-9}, gray on={10-}}, edge label={node[pos=0.6, above right] {$6$}}
      [\varN{4}, edge={alert on={15,26-27}, gray on={16-20,28-}}, edge label={node[midway, right] {$3$}}
        [\varN{3}, edge={alert on={17,29-33}, gray on={18-20,34-}}, edge label={node[midway, right] {$2$}}]]]
  ]
\end{forest}
\end{column}
\hfill
\begin{column}{0.65\textwidth}
  \begin{forest}
    for tree={
      math content,
      grow'=north,
      edge path={},
      inner sep=0pt,
      s=0.2cm,
      s sep=0.6cm,
      l=0.5cm,
      l sep=0.3cm
    }
    [{\varN{1} \thickapprox \varN{5}}, visalert on={2}, name=15
    [{\varN{1} \thickapprox \varN{2}}, visalert on={6}, name=12] 
      [{\varN{2} \thickapprox \varN{3}}, visalert on={5}, name=23
        [{\varN{3} \thickapprox \varN{2}}, visalert on={4}, name=32]
      ]
      [{\varN{3} \thickapprox \varN{5}}, visalert on={7}, name=35
        [{\varN{3} \thickapprox \varN{4}}, visalert on={28}, name=34]
        [{\varN{4} \thickapprox \varN{5}}, visalert on={27}, name=45]
        [{\varN{5} \thickapprox \varN{5}}, visalert on={29}, name=55]
      ]
    ]
    \path let \p1 = (12.south west), \p2 = (15.north) in coordinate (l1l) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (35.south east), \p2 = (15.north) in coordinate (l1r) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (32.south west), \p2 = (23.north) in coordinate (l2l) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (32.south east), \p2 = (23.north) in coordinate (l2r) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (34.south west), \p2 = (35.north) in coordinate (l3l) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (55.south east), \p2 = (35.north) in coordinate (l3r) at (\x1, {(\y1+\y2)/2});
    \path let \p1 = (32.south), \p2 = (12.north west) in coordinate (l4l) at (\x2, {(\y1+\y2)/2});
    \path let \p1 = (32.south), \p2 = (12.north east) in coordinate (l4r) at (\x2, {(\y1+\y2)/2});
    \begin{scope}[thick]
      \draw[visible on=<3->, alert on={3-8}] (l1l) -- (l1r);
      \draw[visible on=<5->] (l2l) -- (l2r);
      \draw[visible on=<21->, alert on={21-29}] (l3l) -- (l3r);
      \draw[visible on=<9->, alert on={9-20}] (l4l) -- coordinate (l4m) (l4r);
      \draw[visible on=<30->, alert on={30-31}] ([yshift=0.1cm]34.north west) -- ([yshift=0.1cm]34.north east);
      \draw[visible on=<32->, alert on={32-35}] ([yshift=0.1cm]55.north west) -- ([yshift=0.1cm]55.north east);
    \end{scope}
  \end{forest}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{The simple algorithm: recap}
\begin{columns}
  \begin{column}{0.3\textwidth}
    \begin{tikzpicture}[scale=1.5, >=Stealth]

  % Clock circle
  \draw[thick] (0,0) circle (1cm);

  % Hour markers
  \foreach \angle in {0,30,...,330} {
    \draw[thick] ({cos(\angle)}, {sin(\angle)}) -- ({0.9*cos(\angle)}, {0.9*sin(\angle)});
  }

  % Clock hands (e.g., 10:10)
  \draw[very thick] (0,0) -- ({0.7*cos(60)}, {0.7*sin(60)}); % Minute hand
  \draw[very thick] (0,0) -- ({0.45*cos(120)}, {0.45*sin(120)}); % Hour hand

  % Counter-clockwise arrows around the clock
  \foreach \angle in {60} {
    \draw[->, thick, TUMAccentOrange, decorate, decoration={snake, amplitude=0.5mm, segment length=2mm, post length=7pt}]
      ({1.2*cos(\angle)}, {1.2*sin(\angle)}) 
      arc[start angle=\angle, end angle=\angle+60, radius=1.2cm];
  }

\end{tikzpicture}
  \end{column}
  \hfill
  \begin{column}{0.6\textwidth}
    \begin{itemize}[<+(1)->]
      \item Recursion on input list 
      \item Independent of tree structure
    \end{itemize}
  \end{column}
\end{columns}
\end{frame}